---
title: "Untitled"
output: html_document
date: "2024-07-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r packages}
#library(ggpubr)
library(ggpmisc)
library(zfit)

library(readxl)
library(knitr)
library(tidyverse)
```

For calculations of Metabolic need according to Ehnes et al, biomass must be in mg, and gives metabolic rates in J/h. Then to prepare the argument for the fluxing function, we need to have biomass in g and metabolic need in J/y. 

# Microarthropods

Individual metabolic need $X$ (J/h) is calculated as $X = M \cdot I$ and : 

$$\ln(I) = \ln(i_P) + a_P \ln(M)+ E_p \cdot \frac{1}{kT}$$

where for Insecta  $\ln(i_P) = 21,97205$, $a_P = 0,758895$, $E_p = 0,6574038$ (Ehnes2011) and $k$ is Boltzmann's constant ($8.62 · 10^{-5}$ eV.K^{-1}).

For epedaphic we have detailed species individual mass, with specific values for adults and juveniles (Chauvat, perso).

```{r EPE, include = F, echo = T}

k = 8.62 * 10^{-5} # eV K-1

EPE <- 
  read_csv("data_tidy/DataMesofaunePerM2.csv") %>% 
  select(where(function(x) !all(is.na(x)))) %>% 
  mutate(across(6:last_col(),
                ~ as.double(.))) %>% 
  select(-c(starts_with("m_"), starts_with("N_"), NbJuv_Ep, PerJuv_Ep, NbJuv_Hem, PerJuv_Hem)) %>% 
  pivot_longer(6:last_col(), names_to="Type", values_to = "n") %>% 
  left_join(read_csv2("SFW_database/IndMassEpi.csv")) %>% #IndMass in µg
  mutate(Ind.FreshMass = 0.001 * Ind.Mass * 3.33) %>% #conversion in mg and in fresh mass
  left_join(read_csv2("SFW_database/MeanTemp.csv")) %>% #MeanTemp is in °C
  mutate(MeanTemp = MeanTemp + 273.15) %>% #conversion in K
  mutate(lnI = 21.97205 + 0.758895*log(Ind.FreshMass) - 0.6574038/(k*MeanTemp), 
         I = exp(lnI), #I in J h-1 (from Ehnes2011)
         nX = exp(lnI) * n # nX is metabolic rate by species type so individual metabolic rate X multiply by Type abundance J h-1
         ) %>% 
  select(c(1:5, Type, nX)) %>% 
  pivot_wider(names_from = Type, values_from = nX) %>% 
  rowwise() %>% 
  mutate(MetabolicNeed.Epe = sum(c_across(6:last_col()), na.rm = T)) %>% # J h-1
  ungroup() %>% 
  #we then add again fresh biomass
  select(c(1:5, MetabolicNeed.Epe)) %>% 
  add_column(Mass.Epe = # mg
               read_csv("data_tidy/DataMesofaunePerM2.csv") %>% 
               select(where(function(x) !all(is.na(x)))) %>%
               select(-c(starts_with("m_"), starts_with("N_"), NbJuv_Ep, PerJuv_Ep, NbJuv_Hem, PerJuv_Hem)) %>% 
               pivot_longer(6:last_col(), names_to="Type", values_to = "n") %>% 
               mutate(n= if_else(is.na(n), 0, n)) %>% 
               left_join(read_csv2("SFW_database/IndMassEpi.csv")) %>% #IndMass in µg
               mutate(FreshMass = 0.001*Ind.Mass * #conversion in mg
                        n * # multiplication by abundance
                        3.33 #conversion to FreshMass
                      ) %>%  
               select(-Ind.Mass, - n) %>% 
               pivot_wider(names_from = Type, values_from = FreshMass)  %>%  
               rowwise() %>% 
               mutate(FreshMass = sum(c_across(6:last_col()))) %>% 
               #select(c(1:5, Mass)) %>% 
               pull(FreshMass))

```

For other micro-arthropods, we use the amount of individual C assessed during 

```{r microarthropods-table, include = F}

microarthr <- 
  read_csv2("data_imports/Bilan_isotopie_meso.csv") %>%
  select(Numero, Bloc, Group, C_amount, N, Weight) %>% 
  left_join(read_csv("data_imports/mesoNcapsA.csv") %>% 
              pivot_longer(cols = starts_with("N_"), 
                           names_to = c("bin", "Group"), 
                           names_sep = "_", 
                           values_to = "N.caps") %>% 
              select(-bin, - Sp), 
            by = c("Numero", "Bloc", "Group")) %>% 
  mutate(N.caps = if_else(is.na(N.caps),
                          N,
                          N.caps), # we join bloc A in N.casp
         m.ind = if_else(Weight > 0,
                         Weight/N.caps,
                         NA)) %>% 
  select(-N) %>% 
  mutate(C.ind = C_amount/N.caps) %>% #individual carbon content 
  mutate(m.ind.C = # individual mass based on carbon content
           if_else(Group == "gam",  #m.ind.C is in µg/ind
                         C.ind/0.413, #where 0.413 is mesostigmata's PropC according to Potapov2022 SupMat
                         C.ind/0.451) #where 0.541 is Collembola's PropC according to Potapov2022 SupMat  
         ) %>% 
  right_join(
    read_csv("data_tidy/DataMesofaunePerM2.csv") %>%
      select(- ends_with(c("_Hem", ".Ad", ".Juv", "_Ep"), ignore.case = FALSE),
             - starts_with("m_tot")) %>%
      pivot_longer(cols = starts_with("N_tot"), 
                   names_to = c("mes", "bin", "Group"), 
                   names_sep = "_", 
                   values_to = "N_tot") %>% 
      select(-c(mes, bin, gf, Sp, mod)) %>% 
      mutate(Group = if_else(Group == "ep",
                             "epi",
                             Group)), 
    by = c("Numero", "Bloc", "Group")) %>%
  mutate(m.tot.C = m.ind.C*N_tot) %>% 
  arrange(Group, Bloc, Numero) %>% 
  filter(Group != "epi")

```



```{r HEMI, echo = T, include = F}

HEM <- 
  microarthr %>% 
  filter(Group== "hem") %>%
  left_join(read_csv2("SFW_database/FWDWRatio.csv")) %>% 
  select(Numero, Bloc, m.ind, m.ind.C, N_tot, FW.over.DW.Ratio) %>% 
  mutate(
    Ind.FreshMass = m.ind.C*0.001* #convertion of µg to mg
                    FW.over.DW.Ratio,
    Ind.FreshMass = if_else(is.na(Ind.FreshMass),
                       mean(Ind.FreshMass, na.rm = TRUE),
                       Ind.FreshMass
                       )) %>% 
  left_join(read_csv2("SFW_database/MeanTemp.csv")) %>% #MeanTemp is in °C
  mutate(MeanTemp = MeanTemp + 273.15) %>% #conversion in K
  mutate(lnI = 21.97205 + 0.758895*log(Ind.FreshMass) - 0.6574038/(k*MeanTemp), #I in J h-1 (from Ehnes2011) 
         I = exp(lnI),
         MetabolicNeed.Hem = exp(lnI) * N_tot, #J/h
         Mass.Hem = Ind.FreshMass*N_tot) %>% #mg
  select(Numero, Bloc, MetabolicNeed.Hem, Mass.Hem) %>%
  group_by(Bloc, Numero) %>% 
  summarise(across(where(is.numeric),
                   mean, na.rm=TRUE)) %>% 
  ungroup()

``` 


For Mesostigmata  $\ln(i_P) = 9.674023$, $a_P = 0.6904864$, $E_p = 0.3792541$ (Ehnes2011).

```{r MESOSTI,echo = T, include = F}

GAM <- 
  microarthr %>% 
  filter(Group== "gam") %>%
  left_join(read_csv2("SFW_database/FWDWRatio.csv")) %>% 
  select(Numero, Bloc, m.ind, m.ind.C, N_tot, FW.over.DW.Ratio) %>% 
  mutate(
    Ind.FreshMass = m.ind.C*0.001* #convertion of µg to mg
      FW.over.DW.Ratio,
    Ind.FreshMass = 
           if_else(
             is.na(Ind.FreshMass),
             mean(Ind.FreshMass, na.rm = TRUE),
             Ind.FreshMass
           )
  ) %>% 
  left_join(read_csv2("SFW_database/MeanTemp.csv")) %>% #MeanTemp is in °C
  mutate(MeanTemp = MeanTemp + 273.15) %>% #conversion in K
  mutate(lnI = 9.674023 + 0.6904864*log(Ind.FreshMass) - 0.3792541/(k*MeanTemp), #I in J h-1 (from Ehnes2011)
         I = exp(lnI),
         MetabolicNeed.Gam = exp(lnI) * N_tot,#J/h
         Mass.Gam = Ind.FreshMass*N_tot) %>% #mg
  select(Numero, Bloc, MetabolicNeed.Gam, Mass.Gam) %>% 
  group_by(Numero, Bloc) %>% 
  summarise(across(where(is.numeric),
                   mean, na.rm=TRUE)) %>% 
  ungroup()
```

```{r tidy-microarthr, echo = T, include = F}

Microarthropods <- 
  EPE %>% 
  left_join(HEM%>%
              group_by(Numero, Bloc) %>% 
              summarise(MetabolicNeed.Hem = mean(MetabolicNeed.Hem),
                        Mass.Hem = median(Mass.Hem)), 
            by = c("Numero", "Bloc")
            ) %>% #because we have a duplicate in Gam that we need to adress 
  left_join(GAM %>%
              group_by(Numero, Bloc) %>% 
              summarise(MetabolicNeed.Gam = mean(MetabolicNeed.Gam),
                        Mass.Gam = median(Mass.Gam)), 
            by = c("Numero", "Bloc") 
             ) %>% #because we have a duplicate in Gam that we need to adress 
  pivot_longer(cols = 6:last_col(), names_to ="Type", values_to = "mes") %>% 
  separate_wider_delim(Type, delim =".", names = c("Type","Group")) %>%
  pivot_wider(names_from = Type, values_from = mes) %>% 
  mutate(MN.J.y.m2 = MetabolicNeed *24*360, #conversion from J/h/m2 to J/y/m2
         FreshMass.g.m2 = Mass*0.001) %>%  #conversion from mg to g
  select(-MetabolicNeed,-Mass) 
  
  
write_csv2(Microarthropods,"SFW_data_tidy/Microarthropods.csv")

Microarthropods %>% 
  group_by(mod, Group) %>% 
  summarise(
    MN.J.y.m2 = mean(MN.J.y.m2),
    MN.kJ.d.m2 = mean(MN.J.y.m2) * 0.001 / 360,
    Mass.mg.m2 = mean(FreshMass.g.m2)*1000)

```

# Nematodes 

Because there is no specific values, we use the generic equation, where $\ln(i_P) = 23.055335$, $a_P = 0.695071$, $E_p = 0.68642$ (Ehnes2011). As for epedaphic, we have more precise data than our trophic group, so metabolic needs are calculated as finely as possible, and then, sumed per group.

```{r nematodes_biomass_per_trophic_group , include = F, echo = T}

dico.nem <- read_csv2("SFW_database/NematodesFamily.csv", col_names = c("Family", "TrophicGroup", "FreshBodyMass"), skip =1)

#bien penser à remultiplier par l'abondance totale pour revenir en nem/gsol

Nematodes <- 
  read_csv2("data_tidy/DataNematodesPerM2.csv") %>% 
  pivot_longer(cols= 9:last_col(), names_to = "Family", values_to = "CountPerMil") %>% 
  inner_join(dico.nem) %>%  #adding trophic group and fresh body mass data [µg]
  left_join(read_csv2("SFW_database/MeanTemp.csv")) %>% #MeanTemp is in °C
  mutate(
    FreshBodyMass = FreshBodyMass * 0.001, #conversion from µg to mg
    MeanTemp = MeanTemp + 273.15, #conversion in K
    FreshMass.fam =  CountPerMil * NemPerM2 * FreshBodyMass, 
    lnI = 23.055335 + 0.695071*log(FreshBodyMass) - 0.68642/(k*(MeanTemp)), #I in J h-1 (from Ehnes2011) 
    X = exp(lnI), # metabolic need per ind for each family
    MetabolicNeed.fam = CountPerMil * NemPerM2 * X # family metabolic need (metabolic need per ind * N ind) - J h-1
         ) %>% 
  #filter(Family == "Tylenchidae")

  
  group_by(Numero, Bloc, Sp, mod, gf, TrophicGroup) %>%
  summarise(MetabolicNeed.TG.nem = sum(MetabolicNeed.fam, na.rm = TRUE),
            MN.TG.nem.J.y.m2 = MetabolicNeed.TG.nem*24*360,  #conversion from J/h/m2 to J/y/m2
            Mass.TG.nem.g.m2 = sum(FreshMass.fam, na.rm = TRUE)*0.001,
            #MN.TG.nem.kJ.d.m2 = sum(MetabolicNeed.fam, na.rm = TRUE) *24 * 0.001
            ) %>% #conversion from mg to g
  select(-MetabolicNeed.TG.nem) %>%
  ungroup()# %>% 
  # group_by(TrophicGroup, mod) %>% 
  # summarise(
  #   MN.TG.nem.J.y.m2 = mean(MN.TG.nem.J.y.m2)
  # )


rm(dico.nem)


# data.nem.biomass %>% 
#   filter(TrophicGroup == "Omnivores" & Mass.TG.nem.m2 > 0)

write_csv2(Nematodes, "SFW_data_tidy/Nematodes.csv")

```

# Microbes

We based our choice of PLFA on Joergensen 2021's classification. We tested a few combinations of conversion factors found in the litterature. 


## Conversion from PLFA concentration to PLFA Biomass

Conversion factors (µgC/nmol PLFA) already take into account the fact that there are un-identified PLFA. We have:


- two value for AMF conversion factor : 366 from Joergensen2008 synthesis but it says it might overestimate by 5-10 times or 26.32 from Olsson1995,


- two solutions for bacteria: either considering the generic factor of 2.75 or taking into account that Gram- bacteria contain ~ 50% more PLFA than G+ bacteria (Joergensen2008)


- two solution for fungal biomass: only taking into account 18:2w6,9 with the conversion factor of 107 (Joergensen2008) or a re-calculation from KlamerBaath2004 of a factor based on the sum of 18:2w6,9 and 18:1w9 of 60.79 (see Excel File conversion fungal PLFA Klamer&Baath for more detail on this recalculation).


We chose scenario 6 because it was the most consistent with our total estimation of microbial biomass according to the fumigation-extraction method. 

```{r include = TRUE, echo= FALSE}

knitr::kable(read_csv2("SFW_database/ConversionFactorC_micro.csv") %>% 
               filter(Group != "unsp_microb"), 
  caption = "Overview of Conversion factors combinaisons")
  
```


There is a significant difference in the dominant fungal group in presence of plants, which might slightly biais our estimation according to modality (underestimating a bit fungal biomass when Zygomycota are more dominant, so in Litter and Control), because conversion factor 6 is based only on 18:2w6,9 and does not take into account 18:1w9.

Microbial C is then converted into microbial biomass according to the table below with C content (µgC/µg): 

```{r ConversionMicrobialBiomass include = TRUE, echo= FALSE}
knitr::kable(read_csv2("SFW_database/ConversionFactorBiomasse.csv") %>% 
  filter(Group != "unsp_microb"),
  caption = "Microbial C content (µgC/µg)")
```



```{r dataPLFAC}

data.PLFA.C <-
  read_csv2("data_imports/DataPLFA.csv") %>% 
  pivot_longer(c(5:last_col()), 
               names_to = "Component",
               values_to = "Concentration") %>% 
  left_join(read_csv2("SFW_database/PLFAGroups_JoergMW.csv")) %>% 
  mutate(Component = if_else(Component ==  "18:1w9, 18:3w3, 18:2w6t",
                             "18:1w9", 
                             if_else(Component == "18:1w7, 18:1w9t",
                                     "18:1w7",
                                     Component)
         )) %>% 
  drop_na(Group) %>% # keeping only groups identified as microbes by Joergensen2021
  filter(Group != "unsp_microb" & Group != "Zygomycota") %>% 
  group_by(Numero, Sp, gf, Group) %>% 
  summarise(Concentration = sum(Concentration)) %>% # PLFA concentration nmol/gsoil for each group
  ungroup() %>% 
  pivot_wider(values_from = Concentration, names_from = Group) %>%
  rowwise() %>% 
  mutate(PLFA_tot = sum(c_across(4:last_col()))) %>% 
  ungroup() %>% 
  pivot_longer(c(4:last_col(), -PLFA_tot), names_to = "Group", values_to = "Concentration") %>% 
  left_join(read_csv2("SFW_database/ConversionFactorC_micro.csv") %>% select(Group, ConvFact6)) %>%
  rename(ai = ConvFact6) %>% # ai [µgC/nmol PLFA]
  mutate(C_micro = Concentration*ai) %>% #conversion from nmol/gsoil to µgC/gsoil
  select(-c(Concentration, ai)) %>% 
  left_join(read_csv2("SFW_database/ConversionFactorBiomasse.csv")) %>%
  mutate(Biomass = 1/ConvFact * C_micro) %>%  #because ConvFact = C_content [µg/gsol]
  select(-c(Type, ConvFact, PLFA_tot))


save("data.PLFA.C", file = "object/data.PLFA.C.Rdata")
```


We calculated proportion of biomass for each group in each modality (PLFA being measured on a composite sample of the three replicates of one modality), which we multiplied by the total microbial C extracted in the fumigation-extraction protocol.


```{r}
data.PLFA.C %>% 
  filter(Numero == 46)
```

```{r microbial.biomass, echo = T, include=FALSE}

microbial.biomass <- # microbial biomass per gSoil
  read_csv2("data_tidy/DataSol.csv") %>% 
  select(Numero, Bloc, Sp, gf, mod, 
         C_micro) %>% # µgC/gsol
  rename(C_micro_FE = C_micro) %>% 
  left_join(data.PLFA.C %>% 
              select(-Biomass) %>% 
              pivot_wider(names_from = Group, values_from = C_micro) %>% 
              rowwise() %>% 
              mutate(C_tot = sum(c_across(4:last_col()))) %>% 
              ungroup() %>%
              pivot_longer(c(4:last_col(), -C_tot), names_to = "Group", values_to = "C_micro") %>% 
              mutate(PropC = C_micro/C_tot) %>% # the contribution of each microbial group to total biomass 
              select(Numero, Group, PropC),
            by = "Numero") %>% 
  mutate(C_micro_group = C_micro_FE*PropC) %>% 
  left_join(read_csv2("SFW_database/ConversionFactorBiomasse.csv")) %>%
  mutate(µgBiomassPerGSoil = 
           1/ConvFact * C_micro_group  #because ConvFact = C_content [µgC/µg_microb]
           ) %>%  
  select(-c(ConvFact))


```




## Metabolic needs and absorption efficiencies

For microbes, metabolic rate is equal to respiration, but we need the energetic equivalent of respirated CO~2~, not the respiration rate in ppm CO~2~.


#### From CO~2~ Respirated to Energy per m^2^ 

##### Objectives

We want the amount of CO~2~, $n_{{CO_2}_{bottle}}$ in the incubation bottle at the end of the incubation period, so we can convert it to energy (we can't do that directly from concentration), using a conversion factor of 1 g C = 39 kJ based on the enthalpy of glucose combustion (Sandler & Orbey, 1991).

**Law of ideal gas:** $PV = nRT$ or, in the case we are interested in $n = PV/RT$.


With the pressure $P$ ($P_{tot}$) in Pa, the volume $V$ in m^3^, the amount of gas substances $n$ in mol (in our case $n_{{CO_2}_{bottle}}$), the temperature $T$ in K, and $R = 8.3145$ the gas constant. The equation is also applicable to gas i at the concentration $x_i$ with $P_i$ the partial pressure and $V_i$ the partial volume, where $x_i = \frac{n_i}{n_{tot}} = \frac{P_i}{P_{tot}} = \frac{V_i}{V_{tot}}$. 


However, we haven't measured the total pressure $P_{tot}$ at the end of the incubation period, only $x_{{CO_2}_{bottle}}$ so we are missing a data to calculate $n_{{CO_2}_{bottle}}$. However, as the bottle were filled with without CO~2~ atmosphere before the beginning of incubation, and we know the stoichiometry of respiration reaction, we can calculate it. 


**Aerobic heterotrophic respiration:** 6 C~6~H~12~O~6~ + 6 O~2~ $\rightarrow$ 6 CO~2~ + 6 H~2~O.

We can deduce that every mole of consumed O~2~ is replaced by a mole of H~2~O, and every CO~2~ mole augments the total pressure.
Thus, 

$P_{tot} = P_{atm} + P_{CO_2}$ and $x_{CO_2} = \frac{P_{CO_2}}{P_{atm} + P_{CO_2}}$, so :

$$P_{CO_2} = \frac{x_{CO_2} \cdot P_{atm}}{1 - x_{CO_2}}$$

$$ n_{CO_2} = \frac{P_{CO_2} · V_{tot}}{R \cdot T} = \frac{P_{tot}· V_{tot}}{R \cdot T}  · x_{CO_2}  $$


```{r PVnRT}
Patm = 1.01325e5 #en Pa
R = 8.3145 # gas constant J. K-1 . mol-1
Q10 = 2

metabolic.microb <- 
  read_csv2("data_imports/IncubTime_CO2.csv") %>% 
  drop_na() %>% 
  mutate(Numero = as.integer(Numero),
         Bloc = as.factor(Bloc),
         xCO2 = XCO2 *10^{-6},#ppm into 0-1 range
         PCO2 = xCO2*Patm/(1-xCO2),
         Ptot = Patm + PCO2,
         nCO2.bottle = # mol
           Ptot * 0.5e-3 * xCO2 / #Ptot en Pa * V = 0.5 L converted in m3
           (R * (25+ 273.15) ) #T_incub = 25 °C converted in K
         ) %>% 
  select(Numero, Bloc, nCO2.bottle, IncubTime) %>% 
  
  mutate(
    gCbottle = #g
      nCO2.bottle *12, # mol * g/mol
    µgCPerGSoilPerH = #µgC-CO2 / g Soil / h
      gCbottle *10^6 / #µgC-CO2
      50 / # G Soil
      (IncubTime /60) # min / min/h
    
  ) %>%  #so far OK
  
  filter(!is.na(Numero)) %>% 
  left_join(microbial.biomass %>% #µgC / gsol
              select(Numero, Bloc, C_micro_FE) %>% 
              unique() %>% 
              rename(µgCmicroPergSoil = C_micro_FE) %>% 
              mutate(Numero = as.integer(Numero),
                     Bloc = as.factor(Bloc)) %>% 
              filter(!is.na(Numero)),
            by = c("Numero", "Bloc")) %>% 
  left_join(read_csv2("SFW_database/MeanTemp.csv")) %>% #MeanTemp is in °C
  mutate(MetRate.µgCO2.µgCmicro.H = #µg C-CO2 / µg C-microb / h
           µgCPerGSoilPerH / # µg C-CO2 / g Soil / h
           µgCmicroPergSoil, # µg C-microb / g soil
         #temperature correction
         qCO2 = MetRate.µgCO2.µgCmicro.H * Q10 ^ ((MeanTemp - 25)/10) #gC-CO2 / gC-microb / h
         ) 

#save("full.M.MN.kJ.d", file = "object/full.M.MN.kJ.d.Rdata")
save("metabolic.microb", file = "object/metabolic.microb.Rdata")

```




#### Metabolic quotients - unplant pots corrected

Everything here is presented with $qCO_2$, which is the metabolic rate expressed as ppm CO~2~ per gram of microbial C. What we have is the metabolic rate expressed as kJ per gram of microbial C, but the reasoning still apply (I will rewrite everything later, but as I started the calculations with $q_CO2$ before noticing I needed the conversion to kJ done at the first step, I haven't rewritten it yet, but you just have to replace $q_CO2$ by "metabolic rate / MN"). 


The metabolic quotient of soil microbes (qCO_2, g C‐CO_2 g-1 Cmicrobial hour-1) is basal respiration divided by microbial biomass, and was standardized to the plot-specific soil temperature using the following equation (Xu et al., 2017):

$$qCO_{2_{pot}} = qCO_{2_{incub}} * Q_{10}^{\frac{T_{field}-T_{incub}}{10}} $$

where $qCO_{2_{pot}}$ and $qCO_{2_{incub}}$ are the $qCO_2$ for field and incubation conditions, respectively, $T_{field}$ and $T_{incub}$ are the bloc specific soil temperature (mean across growth period) and the incubation temperature (25 °C), respectively. $Q_{10}$ is the temperature sensitivity coefficient, and a $Q_{10}$ value of 2 was used (Xu et al., 2017). 

Assuming that bacteria breath twice as much as fungi (SakamotoOba1994) :

$$qCO_{2_{bact}} =  2 \cdot qC0_{2_{fungi}}$$


As we also have

$$qCO_{2_{fungi}} = qCO_2 \cdot  x_{fungi} $$

and the equivalent for bacteria, we can deduce that :

$$qCO_{2_{fungi}} = \frac{qCO_2}{2 \cdot x_{bacteria} + x_{fungi}}$$

where $x_{bacteria}$ and $x_{fungi}$ are respectively the biomass fraction of each group.

Assuming that there is no difference of respiration rate within bacterial and fungal groups, we also have (where $x_{_subgroup}$ = $\frac{subgroup}{group}$) :


$$ qCO_{2_{Gneg}} = qCO_{2_{bact}} \cdot x_{Gneg} $$
$$ qCO_{2_{firmicutes}} = qCO_{2_{bact}} \cdot x_{firmicutes} $$
$$ qCO_{2_{saproF}} = qCO_{2_{fungi}} \cdot x_{saproF} $$
$$ qCO_{2_{AMF}} = qCO_{2_{fungi}} \cdot x_{AMF} $$
However, because AMF are probably in a dormant state or at least activity-diminished state when no plants are present, we will assume their contribution is either null or 10% of its weighted equivalent, so :

$$ qCO_{2_{AMF}} = qCO_{2_{fungi}} \cdot 0.001 \cdot x_{AMF} $$

$$ qCO_{2_{saproF}} = qCO_{2_{fungi}} \cdot \frac{1 - 0.001 \cdot x_{AMF}}{x_{fungi}}  $$

Metabolic need is then calculated as $MN = qCO_{2_{group}} * \text{biomass}_{{group}}$. 



```{r metabolic.quotients-0.001AMF}  

ConvFactGlu = 39 #kJ/gC

Microbes <- 
  microbial.biomass %>% 
  select(1:5, Group, C_micro_group) %>% 
  drop_na(Numero) %>% 
  pivot_wider(values_from = C_micro_group, # µgC micro / g soil
              names_from = Group) %>% 
  rename(Gpos = firmicutes,
         SaproF = AscoBasio) %>% 
  

  mutate(Numero = as.character(Numero)) %>% 
  left_join(read_csv2("SFW_database/MeanTemp.csv")) %>% #MeanTemp is in °C
  left_join(metabolic.microb %>% 
              select(Numero, Bloc, qCO2) %>% 
              mutate(Numero = as.character(Numero))) %>% 
  mutate(TotBiomass = sum(AMF, SaproF, Gneg, Gpos),
         totBact = Gneg+Gpos,
         totFungi = AMF+SaproF,
         xBact = totBact/TotBiomass,
         xFungi = totFungi/TotBiomass,
         qCO2f = qCO2 / (2*xBact + xFungi) , 
         qCO2b = 2*qCO2f, 
         #with the hypothesis that qCO2Gneg = qCO2firmi, qCO2Gneg = qCO2bact * xGneg & idem for fungal groups
         qCO2.Gneg = qCO2b * Gneg/totBact,  
         qCO2.Gpos = qCO2b * Gpos/totBact,
         qCO2.AMF = 
           if_else(
             condition = mod %in% c("P", "P+L"), 
             true = qCO2f * AMF/totFungi,
             false = qCO2f * 0.001*AMF/totFungi
         ),
         qCO2.SaproF = 
           if_else(
             condition = mod %in% c("P", "P+L"), 
             true = qCO2f * SaproF/totFungi,
             false = qCO2f * (1 - 0.001*AMF/totFungi)
         ),
         #Resp|MetabolicNeed group = qCO2 group (# µg C-CO2 / g Soil / h) * biomass group (µg C micro / g soil)
         MN.Gneg = qCO2.Gneg *Gneg, 
         MN.Gpos = qCO2.Gpos * Gpos,
         MN.AMF = qCO2.AMF*AMF,
         MN.SaproF = qCO2.SaproF * SaproF) %>% 
  select(- c(starts_with(c("x", "q", "tot")), MeanTemp)) %>% 
  rename_with(~ if_else(. %in% c("AMF", "SaproF", "Gneg", "Gpos"), 
                        paste0( "biom.", .), 
                        .)) %>% 
  pivot_longer(cols = c(6:last_col()), names_to = "names", 
               values_to = "values") %>% 
  separate_wider_delim(cols=names, delim =".", names=c ("Type", "Group")) %>% 
  pivot_wider(names_from = Type, values_from = values) %>% 
  
     
  ###
  left_join(read_csv2("SFW_database/FWDWRatio.csv") %>% 
              select(Group, FW.over.DW.Ratio) %>% 
              mutate(Group = if_else(Group == "firmicutes", 
                                     "Gpos", Group))
            ) %>% 
  mutate(
    MetabolicNeed.J.y.gsoil =  
           MN * 10^{-6} * # µg C-CO2 --> g C-CO2 
           ConvFactGlu * # conversion from CO2 to kJ with conversion factor = 39 kJ / gC-CO2
           24 * 365 * # hour --> year
           1000, # kJ --> J
         µgBiom.gsoil = biom * FW.over.DW.Ratio) %>% 
  select(-MN, - biom, - FW.over.DW.Ratio) 
  
```


```{r microbial.biomass.g.m2}

Microbes.m2 <- 
  Microbes %>% 
  mutate(gBiomPerM2 = #g Biomass Per m2
           µgBiom.gsoil * #in µg/gsol
           0.8 * #density in g/cm3
           7 * #7 cm3/cm2 (7 cm depth bilan)
           10000 * # conversion from cm2 to square meter
           0.000001,# conversion from µg to g
         MN.perM2 = #J / y / m2
           MetabolicNeed.J.y.gsoil*
           0.8 * #density in g/cm3
           7 * #7 cm3/cm2 (7 cm depth bilan)
           10000  # conversion from cm2 to square meter
  ) %>%  
  select(-c(MetabolicNeed.J.y.gsoil, µgBiom.gsoil)) 


write_csv2(Microbes.m2, "SFW_data_tidy/Microbes_001AMF.csv")
    
```


```{r}
rm(Microbes, EPE, GAM, HEM, lm.B.hem, microarthr,
   metabolic.microb, 
   data.PLFA.C)
```

# Absorption efficiencies and basal resources

Absorption efficiencies are based on this formula (Jochum2017): 
$$\text{logit}(e_a) = 0.471 \cdot \text{FoodN} - 2.097 $$
that is,

$$e_a = \frac{\exp(0.471 \cdot \text{FoodN} - 2.097)}{1 + \exp(0.471 \cdot \text{FoodN} - 2.097)} $$
## Microbes and fauna

For microbes and micro/mesofauna we'll use generic N content from the litterature (Potapov2022).

```{r absorptionEffFauna, echo=T, include=F}

read_csv2("SFW_database/GroupsN_content.csv") %>% 
  mutate(ea = exp(0.471 * N.content - 2.097)/ (1 + exp(0.471 * N.content - 2.097))) %>% 
  mutate(Group = case_when(
                Group == "AMF" ~ "F.amf",
                Group == "SaproF" ~ "F.s",
                Group == "Gneg" ~ "B.Gneg",
                Group == "firmicutes" ~ "B.Gpos",
                Group == "Bacterivores" ~ "Ne.B",
                Group == "Facultatives phyophages" ~ "Ne.fPh",
                Group == "Fungivores" ~ "Ne.F",
                Group == "Phytoparasites" ~ "Ne.PhP",
                Group == "Omnivores" ~ "Ne.Omn",
                Group == "Predators" ~ "Ne.P",
                Group == "Epe" ~ "Cla.Epe",
                Group == "Hem" ~ "Cla.Hem",
                Group == "Gam" ~ "Me")) %>% 
  write_csv2("SFW_data_tidy/AbsorptionEffFauna.csv") 

```


## Basal resources

## Absorption efficiencies

Absorption efficiencies for plant derived resources are calculated from measured N content with the same formula as above. We separated leaf and root litter because their characteristics differ.

## Biomass 

1. We chose to estimate fine root biomass per square meter from root density measured from 14cm-deep core sample. The consequence of this choice we made is that we underestimate a bit the roots quantity for species with an important tap root (**Rumex acetosa*, *Medicago sativa* but also to a lesser extend maybe *Nardus stricta* and *Holcus lanatus* - we don't know why, see file "sandbox - Roots and Rhizodeposition from more infos).

2. Resources for micro-organisms were estimated based on the relative enrichement of DOM in ^15^N for litter and ^13^C for rhizodeposition. 

3. Litter biomass (available for micro-arthropods but not micro-organisms) was calculated from inputs at the begining of the experiment (0.73 g of root litter and 1.47 g of leaf litter in each pot).

```{r basal.resources}
d = 0.8 #soil volumetric  mass density kg/L
V.coresample = #L because 1 dm3 = 1L
  (0.25^2 * pi *1.4) * #length in dm
  3 #3 core samples

#efficiences
BasalsEff <- 
  read_csv2("data_tidy/DataPES.csv") %>% 
  
  ## plants
  select(Numero, Sp, gf, mod, Bloc, Root_N) %>% 
  #we estimate absorption efficiency from fine roots N_content as they are the most likely to be consumed
  bind_rows(
    tibble(Numero = rep(c(31:45), 3),
       Sp = rep(c("N.s." , # 31
         "F.o." , # 32
         "A.o." , # 33
         "D.g." , # 34
         "H.l." , # 35
         "G.v." , # 36
         "C.r." , # 37
         "P.l." , # 38
         "C.n." , # 39
         "R.a." , # 40
         "M.a." , # 41
         "T.p." , # 42
         "L.c." , # 43
         "M.s." , # 44
         "T.r." ), 3), # 45 
       Bloc = c(rep("A", 15), rep("B",15), rep("C",15)),
       mod = rep("L", 45),
       gf = c(rep(c(rep("Grass", 5), rep("Forbs", 5), rep("Leg", 5)), 3)),
       Root_N = rep(NA, 45))) %>% 
  rename(Root_N.FR = Root_N) %>% 
  
  ##litter
  left_join(
    read_csv("data_tidy/DataIsoLit.csv") %>% 
              select(Sp, Type, N_amount) %>%
              pivot_wider(names_from = Type, 
                          names_prefix = "Root_N.", 
                          values_from = N_amount)) %>% 
  pivot_longer(c(starts_with("Root_N")), 
               names_to = c("Group"),
               values_to = "Root_N") %>% 
  mutate(Root_N = if_else(Numero < 16 & 
                            (Group %in% c("Root_N.PS", "Root_N.PA")),
                          NA,
                          Root_N)) %>% 
  drop_na(Root_N) %>% 
  mutate(Group = str_remove(Group, "Root_N."),
         AbsorptionEff = exp(0.471 * Root_N/100 - 2.097)/ (1 + exp(0.471 * Root_N/100 - 2.097))) %>%
  mutate(Group = case_when(
    Group == "FR" ~ "P.FR",
    Group == "PS" ~ "P.rL",
    Group == "PA" ~ "P.lL"
  )) %>% 
  
  #Biomass

  ## adding plant masses converted to g/m2
  left_join(
    read_csv2("data_tidy/PlantBiomasses.csv") %>%
      mutate(Biomass = #g/m2
               R_Dmass_carottes/V.coresample * #root density #g/L = g/dm3
               7 / #depth
               1000 * #cm3/cm2 / 1000 = dm3/cm2 = L/cm2
               10000,  # g/cm2 *10 000 = g/m2
             Group = "P.FR"
      ) %>%
      select(Numero, Bloc, Group, Biomass)
  ) %>%
  select(Numero, Bloc, Sp, mod, gf, Group, AbsorptionEff, Biomass) %>%
  
  ##adding litter masses converted to m2
  
  mutate(Group = as.factor(Group),
         Biomass =
           case_when(
             Group == "P.rL" ~ 0.73 / #g what we put per pot 
               (pi*11.25^2)* #cm2 (11.25 is the mean radius of the pot)
                 10000, # g/cm2 *10 000 = g/m2
             Group == "P.lL" ~ 1.47 / #g what we put per pot
               (pi*11.25^2)* #cm2 (11.25 is the mean radius of the pot)
               10000, # g/cm2 *10 000 = g/m2
               .default =  Biomass
             ))

write_csv2(BasalsEff, "SFW_data_tidy/BasalResources2.csv")



BasalsEff%>%
  group_by(mod, Group) %>%
  summarise(mean.Biomass = mean(Biomass, na.rm = T),
            max = max(Biomass, na.rm = T),
            min = min(Biomass, na.rm = T)) %>%
  arrange(mod, Group)
```


# Completion of input table

```{r full.M.MN}

full.M.MN <- 
  Microbes.m2 %>% 
  mutate(Numero = as.double(Numero),
         Group = case_when(
           Group == "AMF" ~ "F.amf",
           Group == "SaproF" ~ "F.s",
           Group == "Gneg" ~ "B.Gneg",
           Group == "Gpos" ~ "B.Gpos"
         )) %>% 
  rename(Biomass = gBiomPerM2, 
         MetabolicNeed = MN.perM2) %>% 
  bind_rows(Microarthropods %>% 
              mutate(Group = if_else(Group == "Gam",
                                     "Me",
                                     str_c("Cla.", Group))) %>% 
              rename(Biomass = FreshMass.g.m2,
                     MetabolicNeed = MN.J.y.m2)) %>% 
  bind_rows(Nematodes %>% 
              mutate(TrophicGroup = case_when(
                TrophicGroup == "Bacterivores" ~ "Ne.B",
                TrophicGroup == "Facultative phytophages" ~ "Ne.fPh",
                TrophicGroup == "Fongivores" ~ "Ne.F",
                TrophicGroup == "Phytoparasites" ~ "Ne.PhP",
                TrophicGroup == "Omnivores" ~ "Ne.Omn",
                TrophicGroup == "Predators" ~ "Ne.P"
              )) %>% 
              rename(MetabolicNeed = MN.TG.nem.J.y.m2 ,
                     Biomass = Mass.TG.nem.g.m2,
                     Group = TrophicGroup)) %>% 
  left_join(read_csv2("SFW_data_tidy/AbsorptionEffFauna.csv") %>% 
              select(Group, ea)) %>% 
  bind_rows(BasalsEff %>% 
              mutate(MetabolicNeed = 0) %>% 
              rename(ea = AbsorptionEff)
              ) %>% 
  mutate(Group = factor(Group, 
                        levels = c("P.FR" , "P.rL" , "P.lL", "P.dL", "P.Rh" , "S", 
                                   "B.Gpos", "B.Gneg", "F.s", "F.amf",
                                   "Ne.PhP", "Ne.fPh", "Ne.B", "Ne.F", "Ne.Omn", "Ne.P", 
                                   "Me", "Cla.Epe", "Cla.Hem"),
                        ordered = TRUE),
         NumeroBloc = factor(str_c(Numero, ".", Bloc)),
         .after = Bloc) %>% 
  arrange(Bloc, Numero, Group)

# allows us to get every modality
expanded.M.MN <- 
  read_csv2("SFW_database/NumeroSpModGF.csv") %>% 
  mutate(Group = NA,
         NumeroBloc = factor(str_c(Numero, ".", Bloc))) %>% 
  mutate(Group = factor(Group, 
                        levels = c("P.FR" , "P.rL" , "P.lL", "P.dL", "P.Rh" , "S", 
                                   "B.Gpos", "B.Gneg", "F.s", "F.amf",
                                   "Ne.PhP", "Ne.fPh", "Ne.B", "Ne.F", "Ne.Omn", "Ne.P", 
                                   "Me", "Cla.Epe", "Cla.Hem"),
                        ordered = TRUE)) %>% 
  #select(Group, NumeroBloc) %>% 
  expand(Group, NumeroBloc)%>% 
  arrange(NumeroBloc) %>% 
  separate(NumeroBloc, into = c("Numero", "Bloc"), remove = FALSE) %>% 
  left_join(read_csv2("SFW_database/NumeroSpModGF.csv") %>% 
              mutate(Numero=as.character(Numero))) %>% 
  drop_na(Group) 
  

full.M.MN.J.y <- 
  expanded.M.MN %>% 
  left_join(full.M.MN %>% 
              select(Numero, Bloc, Group, MetabolicNeed, Biomass, ea) %>% 
              mutate(Numero = as.character(Numero),
                     Group = as.character(Group))) %>% 
  mutate(across(where(is.numeric),
                ~ if_else(is.na(MetabolicNeed) &
                   is.na(Biomass),
                 0, 
                 .)),
         Biomass = if_else(Group == "S" |
                             (Group == "P.dL") & mod %in% c("Pl+L", "L") |
                              (Group == "P.Rh") & mod %in% c("Pl", "Pl+L"),
                           1, Biomass),
         ea = if_else(Group %in% c("S", "P.dL", "P.Rh"), 1, ea)
         ) %>% 
  filter(Biomass > 0)


full.M.MN.kJ.d <- 
  expanded.M.MN %>% 
  left_join(full.M.MN %>% 
              select(Numero, Bloc, Group, MetabolicNeed, Biomass, ea) %>% 
              mutate(Numero = as.character(Numero),
                     Group = as.character(Group))) %>% 
  mutate(across(where(is.numeric),
                ~ if_else(is.na(MetabolicNeed) &
                   is.na(Biomass),
                 0, 
                 .)),
         Biomass = if_else(Group == "S" |
                             (Group == "P.dL") & mod %in% c("Pl+L", "L") |
                              (Group == "P.Rh") & mod %in% c("Pl", "Pl+L"),
                           1, Biomass),
         ea = if_else(Group %in% c("S", "P.dL", "P.Rh"), 1, ea)
         ) %>% 
  mutate(MetabolicNeed = MetabolicNeed /360 *10^{-3}) %>% 
  filter(Biomass > 0)

rm(expanded.M.MN)


levels(as.factor(full.M.MN.kJ.d$Group))
```




```{r}
#save("full.M.MN.J.y", file = "object/full.M.MN.J.y.Rdata")
save("full.M.MN.kJ.d", file = "object/full.M.MN.kJ.d.Rdata")
```









