---
title: "Fluxing I & II"
output: html_document
date: "2025-10-10"
---

```{r packages}
library(readxl)
library(knitr)
library(tidyverse)
```


```{r load-and-source}
source("FUN_fluxes_calculation.R")
load("object/full.M.MN.kJ.d.Rdata")
```


# Determining microbes/basal interactions factors --------------------------------------------

## Pre-fluxing to determine the losses for each microbial node

```{r output-PRE-fluxing, message=F, warning = F}

#in this matrix, bacteria and fungi are considered as basal resources (no consumption )
mat.pred.as.line.apriori <- read_excel("SFW_data_tidy/Matrix.xlsx", sheet = "a priori")

full.M.MN.ap <- 
  full.M.MN.kJ.d %>% 
  filter(! Group %in% c("P.Rh", "P.dL", "S")) %>% 
  mutate(MetabolicNeed = 
           if_else(Group %in% c("B.Gpos", "B.Gneg", "F.s", "F.amf"),
                   true = 0,
                   false = MetabolicNeed)
         )

mat.fluxes.ap <- mat.fluxes_calculation(full.M.MN.ap, 
                                        mat.pred.as.line.apriori,
                                        microbes.as.consumers = FALSE)

complete_fluxes.ap <- complete_fluxes_calculation(mat.fluxes.ap)

```


## Overview

We did not determine microbial preferences but rather directly the interaction, from the labelling. 13C pulse labelling was used to characterise the relative preference of rhizodeposits ($Rh$) compared to resources from the detritus channel ($D$), litter ($L$) and SOM ($S$).

By definition, 

$$
f_{RhD} = \frac{Rh}{D} \\
f_{SL} = \frac{S}{L} \\
D = S + L
$$
Thus, 

$$
f_{Rh/S} = f_{Rh/D} + \frac{f_{Rh/D}}{f_{S/L}} \\
f_{Rh/L} = f_{Rh/D} + f_{Rh/D} \times f_{S/L}
$$
### Determination of RhDr

The relative consumption factor of rhizodeposits compared to resources from the detritus channel $f_{RhD}$, was determined under two hypothesis :

- fluxes to AMF are 99,999999% rhizofluxes

- the partition of the 13C label across microbial group is a good proxy for the partition of the rhizoflux across microbial groups

#### The partition of 13C label across microbial groups

We calculated enrichment for each microbial group as the mean of the APE of each group's PLFAs, weighted by their concentrations, considering the number of carbon content of each PLFA ($n_{C_{PLFA_i}}$). 

$$
APE_{group} \times Concentration_{group} = \sum_{PLFA_{group}} APE_{PLFA_i} \times Concentration_{PLFA_i} \times n_{C_{PLFA_i}}
$$

where $Concentration_{group} = \sum_{PLFA_{group}} Concentration_{PLFA_i}$.

So, 

$$
APE_{group} = \frac{\sum_{PLFA_{group}} APE_{PLFA_i} \times Concentration_{PLFA_i} \times n_{C_{PLFA_i}}}{\sum_{PLFA_{group}} Concentration_{PLFA_i} \times n_{C_{PLFA_i}}} 
$$


```{r data.APE.µGroups}

data.APE.µGroups <- 
  read.csv("data_tidy/DataIsoPLFA.csv") %>%
  separate(col = Component, sep = ":", into = c("Carb", "Insat"),remove = FALSE) %>% 
  mutate(Nb_Carb = as.numeric(str_extract(Carb, "\\d+")),
         Numero = as.character(Numero)) %>% 
  select(1:5, Group, APE13Cppm, Nb_Carb) %>% 
  drop_na(Group) %>% 
  mutate(APE13Cppm = if_else( APE13Cppm < 0, NA, APE13Cppm)) %>% 
  filter(mod %in% c("P", "P+L") & Group != "unsp_microb") %>% 
  filter(Component != "cy19:0") %>%  #only one value
  group_by(Component) %>%  # Group by the Component variable
  #summarise(APE13Cppm = mean(APE13Cppm, na.rm = T))
  mutate(APE13Cppm = if_else(is.na(APE13Cppm), mean(APE13Cppm, na.rm = TRUE), APE13Cppm)) %>% 
  ungroup() %>% 
  left_join(
    read_csv2("data_imports/DataPLFA.csv")%>% 
      rename(`18:1w9` =  `18:1w9, 18:3w3, 18:2w6t`,
             `18:1w7` = `18:1w7, 18:1w9t`
      ) %>% 
      select(-gf) %>% 
      pivot_longer(cols = 4:last_col(), names_to = "Component", values_to = "Concentration") %>% 
      mutate(Numero = as.character(Numero))
  ) %>% 
  mutate(APE13CppmXconcXnC = APE13Cppm * Concentration * Nb_Carb) %>% #mol de 13C-PLFA = %13C/C * mol-PLFA * molC/molPLFA 
  group_by(Numero, Sp, mod, Group)  %>% 
  mutate(APE13Cppm_m = 
           sum(APE13CppmXconcXnC) /
           sum(Concentration* Nb_Carb)
  ) %>% 
  ungroup()# %>% 
# group_by(Group, mod) %>% 
# summarise(APE13Cppm_m = mean(APE13Cppm_m))

```

```{r data.prop.label.µGroups}
data.prop.label.µGroups <- 
  full.M.MN.kJ.d %>%
  filter(Group %in% c("B.Gpos", "B.Gneg", "F.s", "F.amf") & mod %in% c("Pl", "Pl+L")) %>%
  select(Group, Numero, mod, Bloc, Biomass) %>% 
  rename(Node = Group) %>% 
  mutate(Group = case_when(
    Node == "B.Gpos" ~ "firmicutes",
    Node == "B.Gneg" ~ "Gneg",
    Node == "F.s" ~ "Zygomycota",
    Node == "F.amf" ~ "AMF"),
    .after = Node) %>% 
  left_join(data.APE.µGroups %>% 
              select(Numero, Group, APE13Cppm_m) %>% 
              unique())%>% 
  mutate(APE13Cppm_m = if_else(is.na(APE13Cppm_m), mean(APE13Cppm_m,  na.rm=T), APE13Cppm_m)) %>% 
  left_join(read_csv2("SFW_database/ConversionFactorBiomasse.csv")) %>% 
  mutate(GroupC = ConvFact * Biomass) %>%  #because ConvFact = C_content [µgC/gBiom]
  mutate(N_C_lab_group = GroupC*APE13Cppm_m) %>% #if we wanted absolute values we should multiply that by 10^6 to have the ppm correctly but we are just interested in prop
  select(Node, Numero, Bloc, mod, N_C_lab_group) %>%
  group_by(Numero, Bloc) %>% 
  mutate(
    PropLabGroup = N_C_lab_group/sum(N_C_lab_group)) %>%
  ungroup() 
```

```{r plot-prop-label-µgroups}

paletteGroups <-
  c("#76d2e3", "#ddf4f8","#098372", "#4f098d", "#ccdaf4", "#769be3",
    "#a01307", "#4a838e","#5974aa", "#DAF7A6" ,"grey", "#cd2416","#a4b97d",
    "#09a306","#8B4513","#8B4513", "#f3da0d", "#8B4513", "#DEB887") %>%
  setNames(levels(as.factor(full.M.MN.kJ.d$Group)))

data.prop.label.µGroups %>% 
  mutate(
    #Node = as.ordered(Node)#,
    Node = factor(Node, levels = c("B.Gpos", "B.Gneg", "F.s", "F.amf"))
  ) %>% 
  ggplot(aes(x = Node, y = PropLabGroup*100))+
  geom_hline(yintercept = 50, color = "grey")+
  geom_boxplot(aes(fill = Node))+
  geom_point()+
  #facet_grid(.~mod)+
  labs(y = "% of 13C recovered in microbial groups", x ="Group")+
  theme(#legend.position = "none",
    axis.text.x = element_text(angle = 90))+
  scale_fill_manual(values = paletteGroups)+
  theme_light()+
  theme(legend.position = "none")
```

#### Estimation of RhDr for each group

```{r F_out.µ}
F_out.µ <- 
  full.M.MN.ap %>% 
  filter(Group %in% c("B.Gpos", "B.Gneg", "F.s", "F.amf")) %>% 
  select(-Biomass, - ea) %>% 
  left_join(
    complete_fluxes.ap %>% 
      filter(From %in% c("B.Gpos", "B.Gneg", "F.s", "F.amf")) %>% 
      group_by(Numero, Bloc, mod, From) %>% 
      summarise(Flux = sum(Flux)) %>% 
      rename(Group = From) %>% 
      select(-mod)
  ) %>% 
  mutate(F_out.µ = MetabolicNeed + Flux) %>% 
  select(-MetabolicNeed, - Flux)
```

```{r calc-RhD}
data.calc.RhD <- 
  F_out.µ %>% 
  left_join(
    data.prop.label.µGroups %>%
      select(-PropLabGroup) %>% 
      pivot_wider(names_from = Node, values_from = N_C_lab_group) %>% 
      mutate(across(where(is.numeric),
                    ~ .x / F.amf)) %>% 
      pivot_longer(cols = 4:7, names_to = "Group", values_to = "r.µAMF")
  ) %>% 
  dplyr::group_by(NumeroBloc) %>% 
  dplyr::mutate(
    Rh_in.AMF = (1 - 1e-9)*F_out.µ[Group == "F.amf"],
    Rh_in = 
      if_else(Group == "F.amf", 
              Rh_in.AMF,
              r.µAMF * Rh_in.AMF),
    PropRh = Rh_in/F_out.µ, 
    Check = any(PropRh > 1, na.rm = TRUE),
    RhDr = if_else(Check, NA, PropRh/(1 - PropRh)),
    RhDr = if_else(Group != "F.amf" & RhDr > 100, NA, RhDr) #it's for the few values of PropRh that are almost equal to one, but have not been removed
  ) %>%
  ungroup()


data.calc.RhD %>% 
  filter(Group == "F.s") %>% 
  select(1:5, RhDr, PropRh) %>% 
  filter(RhDr > 10)

data.calc.RhD %>% 
  filter(Group == "F.s") %>% 
  mutate(point_col = if_else(RhDr > 10, "red", "black")) %>% 
  pivot_longer(cols = c(F_out.µ, r.µAMF, Rh_in, Rh_in.AMF, PropRh), names_to = "var") %>% 
  drop_na() %>% 
  ggplot(aes(x = mod, y = value))+
  facet_wrap(~var, scales = "free")+
  geom_boxplot()+
  geom_point(aes(color = point_col))+
  scale_color_manual(values = c("black" = "black", "red" = "red"))
```

####Estimation of f.LS for each pot

```{r data-ape-micro, include = F}
data.APE.micro <-   read_csv2("data_tidy/DataSol.csv") %>% 
  filter(! mod %in% c("P", "T")) %>% 
  select(1:5, N_F, N_NF) %>% 
  pivot_longer(cols= c(N_F, N_NF), names_to = "Type", values_to = "concentration") %>% 
  separate_wider_delim(cols = Type, delim = "_", names = c("Element", "Type")) %>% 
  left_join(
    read_csv("data_tidy/DataAPESoilExtracts.csv") %>% 
      select(1:5, APE15Nppm) 
  ) %>% 
  pivot_wider(names_from = Type, values_from = c(concentration, APE15Nppm)) %>% 
  #unnest()
  mutate(across(where(is.list), ~ sapply(.x, function(x) mean(unlist(x), na.rm = TRUE)))) %>% 
  mutate(concentration_µ = concentration_F - concentration_NF,
         APE15Nppm_µ = 
           (APE15Nppm_F * 10^{-6} * concentration_F - 
              APE15Nppm_NF * 10^{-6} * concentration_NF) / concentration_µ * 10^6
  ) %>% 
  select(1:5,APE15Nppm_µ)

data.ape.lit <- 
  read_csv("data_tidy/DataAPELit.csv") %>% 
  mutate(CN= CperMS/NperMS) %>% 
  select(Sp, Type, CperMS, NperMS, APE15Npermil) %>% 
  mutate(APE15NpermilG = APE15Npermil/1000*15/ (APE15Npermil/1000*15 + (1- APE15Npermil/1000)*14) * 1000) %>%  # conversion from APE permil (mol) to APE permil (mass)
  select(-APE15Npermil) %>% 
  pivot_wider(names_from = Type, values_from = c(CperMS, NperMS, APE15NpermilG)) %>% 
  #pour bien faire est-ce qu'il ne faudrait pas aussi pondérer l'APE par la quanté de NperMS?
  mutate(APE15Npermil= APE15NpermilG_PS*(0.75/(0.75+1.47)) + # gN/gMS * ‰gN * g 
           APE15NpermilG_PA*(1.47/(0.75+1.47)),
         CperMS = CperMS_PS *(0.75/(0.75+1.47)) + 
           CperMS_PA*(1.47/(0.75+1.47)),
         NperMS = NperMS_PS *(0.75/(0.75+1.47)) + 
           NperMS_PA*(1.47/(0.75+1.47))
  ) %>% 
  select(-ends_with(c("_PA","_PS")))

```

**Hypothesis**

Relative microbial preference for SOM and litter was assessed from ^15^N enrichment of microbial biomass, considered to reflect microbial consumption of both sources, under the following hypotheses: 

- Nitrogen decomposed from both litter and SOM sources is completely absorbed by microbes
- CUE is identical for both sources

With:

- $L$ the proportion of flux coming from litter,
- $S$ the proportion of flux coming from SOM,
- $N_{lit_{µ}}$, the proportion of nitrogen in the microbial biomass coming from litter,
- $N_{SOM_{µ}}$, the proportion of nitrogen in the microbial biomass coming from SOM,
- $N_{lit}$, the proportion of nitrogen in the litter,
- $C_{lit}$, the proportion of carbon in the litter,
- $N_{SOM}$, the proportion of nitrogen in SOM,
- $C_{SOM}$, the proportion of carbon in SOM,

Then, we have:

$$
f_{L:S} = \frac{C_{conso_{lit}}}{C_{conso_{SOM}}} 
        = \frac{N_{conso_{lit}} \times \text{C:N}_{lit}}{N_{conso_{SOM}} \times \text{C:N}_{SOM}}
        = \frac{N_{µ_{lit}}/NUE_{lit} \times \text{C:N}_{lit} }
               {N_{µ_{SOM}}/NUE_{SOM} \times \text{C:N}_{SOM}} \\
               
                
N_{µ_{lit}} = \frac{\text{APE}_{N_µ}}{\text{APE}_{N_{litter}}} \\
N_{µ_{SOM}} = 1 - N_{µ_{lit}}

$$

```{r calc-LS}
NUE_lit = 0.9
NUE_som = 0.6 

data.calc.LS <- 
  read_csv2("data_tidy/DataSol.csv") %>% 
  #colnames()
  select(1:7, 13:16) %>% 
  filter(! mod %in% c("P", "T")) %>%  
  left_join(data.ape.lit %>% 
              mutate(CNr_lit = CperMS/NperMS) %>%
              rename(APE15Npermil_lit = APE15Npermil) %>%  
              select(Sp, APE15Npermil_lit, CNr_lit)) %>%  
  left_join(data.APE.micro %>% 
              left_join(data.APE.micro %>% 
                          group_by(Numero) %>% 
                          summarise(APE15Nppm_µ_moy = mean(APE15Nppm_µ, na.rm =T))
              ) %>% 
              mutate(APE15Nppm_µ = if_else(is.nan(APE15Nppm_µ),
                                           APE15Nppm_µ_moy,
                                           APE15Nppm_µ)
              ) %>% 
              select(-APE15Nppm_µ_moy)
  ) %>% 
  left_join(read_csv("data_tidy/DataIsoSol.csv") %>% 
              mutate(CNr_soil = CperMS/NperMS) %>% 
              select(Numero, Bloc, CNr_soil)) %>% 
  mutate(CNr_µ = C_micro/N_micro,
         N.µ.lit = APE15Nppm_µ * 10^{-6} / (APE15Npermil_lit*10^{-3}),
         N.µ.som = 1 - N.µ.lit,
         N.conso.lit = N.µ.lit / NUE_lit,
         N.conso.som = N.µ.som / NUE_som,
         LSr = N.conso.lit* CNr_lit /
                (N.conso.som * CNr_soil),
         SLr = 1/LSr,
         NumeroBloc = str_c(Numero, ".", Bloc)) %>% 
  select(NumeroBloc, mod, LSr, SLr) 

data.calc.LS %>% 
  group_by(mod) %>% 
  summarise(LS.mean = mean(LSr))


```

#### Final ratio table


```{r data-calc}
data.calc <-
  full.M.MN.kJ.d %>% 
  select(1:7) %>% 
  filter(Group %in% c("B.Gpos", "B.Gneg", "F.s", "F.amf")) %>% 
  left_join(data.calc.LS %>% select(-mod)) %>% 
  left_join(data.calc.RhD %>% 
              select(Numero, Bloc, Group, RhDr) 
            ) %>% 
  mutate(
    RhSr = case_when(
            mod == "Pl" ~  RhDr,
            mod == "Pl+L" ~ RhDr+RhDr/SLr, 
            .default = NA),
    RhLr = if_else(mod == "Pl+L",
                  true = RhDr*SLr+RhDr,
                  false = NA)
  ) %>% 
  arrange(as.numeric(Numero), Bloc) %>% 
  select(-Numero, - Bloc) %>% 
  select(-RhDr) 

#gestion des NAs  

data.calc.comp <- 
  data.calc %>%
    pivot_longer(cols = 6:last_col(), names_to = "ratio") %>% 
    left_join(
      data.calc %>% 
        pivot_longer(cols = 6:last_col(), names_to = "ratio") %>% 
        dplyr::group_by(Group, mod, gf, ratio) %>% 
        dplyr::summarise(
          Median = median(value, na.rm = T)) %>% 
        drop_na()
        ) %>% 
    mutate(
      value = 
        case_when(
          is.na(value) & str_detect(mod, pattern = "L") & str_detect(ratio, pattern = "L") |
           is.na(value) & str_detect(mod, pattern = "Pl") & str_detect(ratio, pattern = "Rh") ~ 
            Median, 
          .default = value
        )
    )  %>% 
  select(-Median) %>% 
  drop_na() %>% 
  pivot_wider(names_from = ratio, values_from = value)
  
save(data.calc.comp, data.calc, file = "object/data.calc.Rdata")
```



# Final Fluxing --------------------------------------------

```{r outputfluxing, warning=F}
mat.pred.as.line <- read_excel("SFW_data_tidy/Matrix.xlsx", sheet = "FINAL")

#full.M.MN <- full.M.MN.kJ.d

mat.fluxes <- mat.fluxes_calculation(full.M.MN.kJ.d, mat.pred.as.line)

functions <- 
  functions_calculation(mat.fluxes) %>% 
    separate_wider_delim(col = names, 
                         names = c("Numero", "Bloc"), 
                         delim = ".") %>% 
    left_join(read.csv2("SFW_database/NumeroSpModGF.csv") %>% 
                mutate(Numero = as.character(Numero))) %>% 
    select(is.character, is.double)

complete_fluxes <- complete_fluxes_calculation(mat.fluxes)

save(full.M.MN.kJ.d, mat.fluxes, functions, complete_fluxes, mat.pred.as.line, file = "object/full.results.Rdata")
```

```{r}
full.M.MN.ap


functions %>% 
  filter(predation == 0 |
           is.na(predation))

functions %>% 
  filter(predation > 0 & !is_extreme(predation)) %>% 
  summarise(min(predation),
            max(predation),
            median(predation))

functions %>% 
  filter(herbivory == 0 & mod %in% c("P", "P+L"))

 functions%>% 
  filter(mod %in% c("P", "P+L") & herbivory > 0 & !is_extreme(herbivory)) %>% 
  summarise(min(herbivory),
            max(herbivory),
            median(herbivory))
```

